% ============================================================================
% Chapter 41: Revision — MS Office and Internet
% ============================================================================
\chapter{Revision --- MS Office and Internet}
\label{ch:revision-office}

\begin{objectives}
  \item Revise key concepts from Parts IV--VII (Word, Excel, PowerPoint, Internet)
  \item Compare features across MS Office applications
  \item Test your knowledge with practical review tasks
  \item Consolidate internet safety and email skills
\end{objectives}

\bytetip[tip]{This is Part 2 of your revision --- covering all your MS Office skills and internet knowledge. If you can answer everything here, you're ready for any test!}

\section{MS Office Quick Comparison}
\label{sec:office-compare}

\begin{table}[H]
\centering
\caption{Comparing the three MS Office applications}
\label{tab:office-compare}
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{l X X X}
\toprule
\rowcolor{HeaderRow}
\textcolor{white}{\textbf{Feature}} & \textcolor{white}{\textbf{MS Word}} & \textcolor{white}{\textbf{MS Excel}} & \textcolor{white}{\textbf{MS PowerPoint}} \\
\midrule
\rowcolor{RowLight} Purpose & Documents & Data \& calculations & Presentations \\
\rowcolor{RowDark} Extension & .docx & .xlsx & .pptx \\
\rowcolor{RowLight} Main area & Page & Worksheet (cells) & Slide \\
\rowcolor{RowDark} Key feature & Text formatting & Formulas \& charts & Animations \\
\rowcolor{RowLight} Best for & Reports, letters & Budgets, marks & Slideshows \\
\bottomrule
\end{tabularx}
\end{table}

\section{MS Word Revision}
\label{sec:word-revision}

\begin{reviewqs}
  \item Name 6 parts of the MS Word window.
  \item What is the difference between Save (\key{Ctrl}+\key{S}) and Save As (\key{F12})?
  \item List 4 font formatting options and their shortcuts.
  \item How do you insert a table, picture, and page number in Word?
  \item What is the difference between Portrait and Landscape orientation?
\end{reviewqs}

\section{MS Excel Revision}
\label{sec:excel-revision}

\begin{reviewqs}
  \item What is a cell address? Give an example.
  \item Write the formulas for: SUM of B1 to B10, AVERAGE of C1 to C5.
  \item What does AutoFill do, and how do you use it?
  \item Name three chart types and when to use each.
  \item How do you sort data in ascending order?
\end{reviewqs}

\section{MS PowerPoint Revision}
\label{sec:ppt-revision}

\begin{reviewqs}
  \item What is the difference between transitions and animations?
  \item Name 4 slide layouts in PowerPoint.
  \item What is the 6×6 rule for slide design?
  \item What keyboard shortcut starts a slideshow?
  \item List 3 golden rules of good slide design.
\end{reviewqs}

\section{Internet \& Safety Revision}
\label{sec:internet-revision}

\begin{reviewqs}
  \item What is the difference between the Internet and the World Wide Web?
  \item Who invented the WWW?
  \item What are the parts of a URL?
  \item Name 4 signs of a phishing email.
  \item What is digital citizenship? List 3 responsibilities.
  \item Explain the CRAAP test for evaluating websites.
\end{reviewqs}

\begin{challenge}
Create a \textbf{one-page cheat sheet} (in Word or by hand) summarising the most important keyboard shortcuts, definitions, and concepts from the entire book. Make it colourful, organised, and something you'd want to stick on your wall!
\end{challenge}
