% ============================================================================
% Chapter 24: Editing Text in MS Word
% ============================================================================
\chapter{Editing Text in MS Word}
\label{ch:word-editing}

\begin{objectives}
  \item Select, copy, cut, paste, and move text efficiently
  \item Use Find and Replace to edit large documents
  \item Undo and redo actions
  \item Use spell check and autocorrect features
\end{objectives}

\bytetip[tip]{Editing is where the real magic happens. You rarely get text perfect on the first try --- and that's fine! Word's editing tools make fixing things super easy.}

\section{Selecting Text}
\label{sec:selecting-text}
\index{MS Word!selecting text}

\begin{theorybox}[Selection Techniques]
\begin{tabular}{ll}
  \textbf{To select\ldots} & \textbf{Do this\ldots} \\
  \midrule
  A word & Double-click the word \\
  A line & Click in the left margin next to the line \\
  A sentence & Hold \key{Ctrl} and click anywhere in the sentence \\
  A paragraph & Triple-click anywhere in the paragraph \\
  All text & Press \key{Ctrl}+\key{A} \\
  Custom range & Click at start, hold \key{Shift}, click at end \\
\end{tabular}
\end{theorybox}

\section{Copy, Cut, and Paste}
\label{sec:copy-cut-paste}
\index{MS Word!copy paste}

\begin{shortcutbox}
\begin{tabular}{lll}
  \key{Ctrl}+\key{C} & Copy & Copies selected text to clipboard \\
  \key{Ctrl}+\key{X} & Cut & Removes text and copies to clipboard \\
  \key{Ctrl}+\key{V} & Paste & Inserts text from clipboard \\
  \key{Ctrl}+\key{Z} & Undo & Reverses the last action \\
  \key{Ctrl}+\key{Y} & Redo & Re-applies an undone action \\
\end{tabular}
\end{shortcutbox}

\section{Find and Replace}
\label{sec:find-replace}
\index{MS Word!Find and Replace}

Need to change ``colour'' to ``color'' throughout a 20-page document? Don't do it manually!

\begin{labstep}[Using Find and Replace]
\begin{enumerate}[label=\protect\stepcircle{\arabic*}, leftmargin=2cm, itemsep=3pt]
  \item Press \key{Ctrl}+\key{H} to open Find and Replace.
  \item In ``Find what,'' type the word you want to find.
  \item In ``Replace with,'' type the new word.
  \item Click ``Replace All'' to change every occurrence at once.
  \item A message tells you how many replacements were made.
\end{enumerate}
\end{labstep}

\section{Spell Check}
\label{sec:spell-check}
\index{MS Word!spell check}

Word automatically underlines spelling errors with a \textcolor{WarnRed}{red wavy line} and grammar errors with a \textcolor{ModuleBlue}{blue wavy line}. Right-click the underlined word to see suggestions.

\bytetip[fun]{Word's spell checker knows over \textbf{100,000 words} in English! But it can't catch everything --- if you type ``their'' when you mean ``there,'' it won't notice because both are real words. Always proofread your work!}

\begin{funfact}
The clipboard (where copied text goes) was named after a real \textbf{clipboard} --- the board with a clip that office workers used to hold papers. The computer version was invented at Xerox PARC in the 1970s.
\end{funfact}

\begin{reviewqs}
  \item How do you select an entire paragraph in Word?
  \item What is the difference between Cut and Copy?
  \item What shortcut opens Find and Replace?
  \item What does a red wavy underline mean in Word?
  \item How do you undo the last action?
\end{reviewqs}

\begin{challenge}
Type a paragraph of at least 6 sentences with \textbf{intentional spelling mistakes}. Use Word's spell checker to find and fix all of them. Then use Find and Replace to change one word throughout the paragraph. Count how many replacements were made.
\end{challenge}
