% ============================================================================
% Chapter 11: Control Panel and Settings
% ============================================================================

\chapter{Control Panel and Settings}
\label{ch:control-panel}

\begin{objectives}
  \item Access the Windows Settings and Control Panel
  \item Change basic system settings (display, date/time, volume)
  \item Understand the purpose of common Control Panel categories
  \item Personalise the computer's appearance
\end{objectives}

\bytetip[tip]{The Control Panel and Settings are like the ``command centre'' of your computer. This is where you tweak, tune, and customise everything!}

% ---------------------------------------------------------------
\section{Settings vs Control Panel}
\label{sec:settings-vs-cp}
\index{Control Panel}\index{Settings}

Windows has two places to change settings:
\begin{itemize}[leftmargin=*, label={\textcolor{ModuleBlue}{\faAngleRight}}, itemsep=3pt]
  \item \textbf{Settings} (modern, touch-friendly) --- press \key{Windows}+\key{I}
  \item \textbf{Control Panel} (classic, more detailed) --- search ``Control Panel'' in Start
\end{itemize}

% --- Figure 11.1: Control Panel Categories Grid ---
\begin{figure}[H]
\centering
\begin{tikzpicture}[
  tile/.style={
    draw=#1, fill=#1!10, rounded corners=6pt,
    minimum width=3cm, minimum height=2cm,
    align=center, font=\small\sffamily, line width=1pt
  }
]
  % Row 1
  \node[tile=ModuleBlue] at (0,0) {{\Large\textcolor{ModuleBlue}{\faDesktop}}\\System};
  \node[tile=FunPurple] at (3.8,0) {{\Large\textcolor{FunPurple}{\faSun}}\\Display};
  \node[tile=NoteOrange] at (7.6,0) {{\Large\textcolor{NoteOrange}{\faClock}}\\Date/Time};
  
  % Row 2
  \node[tile=LabGreen] at (0,-2.8) {{\Large\textcolor{LabGreen}{\faPrint}}\\Printers};
  \node[tile=FunSky] at (3.8,-2.8) {{\Large\textcolor{FunSky}{\faNetworkWired}}\\Network};
  \node[tile=FunCoral] at (7.6,-2.8) {{\Large\textcolor{FunCoral}{\faUserShield}}\\Accounts};
  
  % Row 3
  \node[tile=WarnRed] at (0,-5.6) {{\Large\textcolor{WarnRed}{\faLock}}\\Security};
  \node[tile=FunMint] at (3.8,-5.6) {{\Large\textcolor{FunMint}{\faVolumeUp}}\\Sound};
  \node[tile=FunPink] at (7.6,-5.6) {{\Large\textcolor{FunPink}{\faCubes}}\\Programs};
\end{tikzpicture}
\caption{Control Panel categories --- your computer's command centre}
\label{fig:cp-grid}
\end{figure}

\section{Common Settings You Can Change}
\label{sec:common-settings}

\begin{enumerate}[label=\textcolor{ModuleBlue}{\textbf{\arabic*.}}, itemsep=4pt]
  \item \textbf{Display Settings} --- Change screen brightness, resolution, and text size
  \item \textbf{Date and Time} --- Set the correct date, time, and time zone
  \item \textbf{Sound Settings} --- Adjust volume, choose notification sounds
  \item \textbf{Personalisation} --- Change wallpaper, colours, lock screen image
  \item \textbf{User Accounts} --- Create accounts, set passwords, manage users
\end{enumerate}

\bytetip[warn]{Be careful when changing settings you don't understand! Some changes (like network or security settings) can stop your internet from working or lock you out. If in doubt, ask your teacher.}

\begin{funfact}
The Control Panel has been part of Windows since \textbf{Windows 3.0} in 1990. Microsoft has been slowly moving features into the newer ``Settings'' app, but the classic Control Panel still exists and is used by IT professionals every day!
\end{funfact}

\begin{reviewqs}
  \item What is the keyboard shortcut to open Windows Settings?
  \item Name four categories found in the Control Panel.
  \item How would you change your desktop wallpaper?
  \item What is the difference between Settings and Control Panel?
  \item Why should you be careful when changing system settings?
\end{reviewqs}

\begin{challenge}
Open \textbf{Settings} $\rightarrow$ \textbf{Personalisation} and change your desktop wallpaper to a solid colour. Then change the system accent colour to match. Finally, change it all back to what it was before!
\end{challenge}
